### 2018-01-05 team.Rmd

---
title: "Lab Menmbers"
---

### 教員

- [Yoshihiko Kunisato](yoshihiko-kunisato.html)

### PD

- 森 正樹　(日本学術振興会 特別研究員PD)

### 大学院生（博士2年）

- [坂本次郎](https://mrunadon.github.io/blogdown_page/post/profile/)　(日本学術振興会 特別研究員DC1)
- [杣取恵太](https://ksomatori.github.io/homepage/index.html)　(日本学術振興会 特別研究員DC1)

### 大学院生（修士２年）

- 金子美帆
- 菊池羽音
- 土谷美月

### 大学院生（修士１年）

- 井口聖香
- 土原浩平

### 学部生（４年）

- 阿久津裕史
- 福地友輔
- 関 海真
- 中川卓知
- 中西由紀奈
- 長尾雅春
- 関谷朋志

### 学部生（３年）
- 広瀬友彩 
- 小嶋那奈 
- アン ミョンヒ 
- 井上圭右 
- 小澤千尋 
- 西田千夏