### 2018-01-05 team.Rmd

---
title: "Team"
---

### Principal investigator

- [Yoshihiko Kunisato](yoshihiko-kunisato.html)

### Postdoctoral Fellow

- Masaki Mori　(JSPS Research Fellowship for Young Scientists, Postdoctoral Fellow (PD))

### Ph.D. students

- [Jiro Sakamoto](https://mrunadon.github.io/blogdown_page/post/profile/)　(JSPS Research Fellowship for Young Scientists, Doctoral Course Students (DC1))
- [Keita Somatori](https://ksomatori.github.io/homepage/index.html)　(JSPS Research Fellowship for Young Scientists, Doctoral Course Students (DC1))

### Master Course students

- Miho Kaneko
- Wato Kikuchi
- Mizuki Tsuchiya
- Seika Iguchi
- Kohei Tsuchihara

### Undergraduate students

- Hirofumi Akutsu
- Yusuke Fukuchi
- Kaima Seki
- Takuto Nakagawa
- Yukina Nakanishi
- Masaharu Nagao
- Tomoyuki Sekiya
- Yua Hirose 
- Nana Kojima 
- Myeonghee Ahn 
- keisuke Inoue 
- Chihiro Ozawa 
- Chika Nishida