### 2018-01-06 index.Rmd
メールアドレスの修正

---
title: "Computational Clinical Psychology Lab"
---

<img src="mark.png" width="600px">

<br />

### Welcome 
Welcome to the Computational Clinical Psychology Lab.
Our research interests are (1)  impairments of decision making in depression using the bayesian cognitive modelling, (2) fear conditioning in anxiety disorders using the neural network model, (3) evedence based practce in clinical psychology.The Computational Clinical Psychology Lab is directed by Yoshihiko Kunisato, Ph.D.

<br />

### Contact
#### Yoshihiko Kunisato, Ph.D.
#### Department of Psychology, Senshu University
#### ADRESS: 2-1-1, Higashimita, Tama-ku, Kawasaki-shi, Kanagawa 214-8580, Japan
#### E-MAIL:ykunisato (at sign) psy.senshu-u.ac.jp