### 2018-04-04 team.Rmd
チームの追加

---
title: "Team"
---

### Principal investigator

- [Yoshihiko Kunisato](yoshihiko-kunisato.html)

### Postdoctoral Fellow

### Ph.D. students

- [Keita Somatori](https://ksomatori.github.io/homepage/index.html)　(JSPS Research Fellowship for Young Scientists, Doctoral Course Students (DC1))

### Master Course students

- Seika Iguchi
- Kohei Tsuchihara
- Takuto Nakagawa

### Undergraduate students

- Hirofumi Akutsu
- Yua Hirose 
- Nana Kojima 
- Myeonghee Ahn 
- keisuke Inoue 
- Chihiro Ozawa 
- Chika Nishida
- Runa Kikuchi
- Mai Shibata
- Momoka Shibaya
- Nodoka Inoue
- Sakina Inoue
- Natsumi Oofusa
- Kaito Suzuki