### 2020-07-23 articles.Rmd
update 23-07-2020

---
title: "Journal Articles"
---

### 2020

- Katahira, K., Kunisato, Y., Okimura, T., & Yamashita, Y. (2020) Retrospective surprise: A computational component for active inference. Journal of Mathematical Psychology, 96, 102347.  [Link](https://www.sciencedirect.com/science/article/abs/pii/S0022249620300286?dgcid=author)

- Katahira, K., Kunisato, Y., Yamashita, Y., & Suzuki, S. (2020) Commentary: "A robust data-driven approach identifies four personality types across four large data sets". Frontiers in Big Data, 3(8). [Link](https://www.frontiersin.org/articles/10.3389/fdata.2020.00008/full)

### 2019

- Nishiguchi, Y., Sakamoto, J., Kunisato, Y., & Takano, K. (2019) Linear Ballistic Accumulator Modeling of Attentional Bias Modification Revealed Disturbed Evidence Accumulation of Negative Information by Explicit Instruction. Frontiers in Psychology,10, 2447.  [Link](https://www.frontiersin.org/articles/10.3389/fpsyg.2019.02447/full)

- Yokomitsu, K., Sakai, T., Irie, T., Tayama, J., Furukawa, H., Himachi, M., Kanazawa, J., Koda, M., Kunisato, Y., Matsuoka, H., Takada, T., Takahashi, F., Takahashi, T., & Osawa, K. (2019). Gambling symptoms, behaviors, and cognitive distortions in Japanese university students. Substance Abuse Treatment, Prevention, and Policy, 14, 51.  [Link](https://substanceabusepolicy.biomedcentral.com/articles/10.1186/s13011-019-0230-5)

- Hasegawa, A., Somatori, K., Nishimura, H., Hattori, Y., & Kunisato, Y. (2019). Associations between self-reported impulsivity and a latent variable of impulsive action constructed from three laboratory tasks. Journal of Experimental Psychopathology, 10, 3. [Link](https://journals.sagepub.com/doi/full/10.1177/2043808719861894)

### 2018

- Hasegawa, A., Kunisato, Y., Morimoto, H., Nishimura, H., & Matsuda, Y. (2018). Depressive rumination and urgency have mutually enhancing relationships but both predict unique variance in future depression: A longitudinal study. Cogent Psychology, 5: 1450919.ß [Link](https://www.tandfonline.com/doi/full/10.1080/23311908.2018.1450919)

- Hasegawa, A., Kunisato, Y., Morimoto, H., Nishimura, H., & Matsuda, Y. (2018). How do rumination and social problem solving intensify depression? A longitudinal study. Journal of Rational-Emotive & Cognitive-Behavior Therapy, 36, 1, 28-46. [Link](https://link.springer.com/article/10.1007/s10942-017-0272-4)

- Takagaki, K., Okamoto, Y., Jinnin, R., Mori, A., Nishiyama, Y., Yamamura, T., Yokoyama, S., Shiota, S., Okamoto, Y., Miyake, Y., Ogata, A., Kunisato, Y., Shimoda, H., Kawakami, N., Furukawa, T. A., & Yamawaki. (2018).  Enduring effects of a five-week behavioral activation program for subthreshold depression among late adolescents: an exploratory randomized controlled trial.  Neuropsychiatric Disease and Treatment , 14, 2633-2641. [Link](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6186299/)

### 2017

- Moriguchi, Y., Noda, T., Nakayashiki, K., Takata, Y., Setoyama, S., Kawasaki, S., Kunisato, Y., Mishima, K., Nakagome, K., Hanakawa, T. (2017). Validation of brain-derived signals in near-infrared spectroscopy through multivoxel analysis of concurrent functional magnetic resonance imaging. Human Brain Mapping, 38(10), 5274–5291.  [Link](https://onlinelibrary.wiley.com/doi/full/10.1002/hbm.23734)

- Yoshimura, S., Okamoto, Y., Matsunaga, M., Onoda, K., Okada, G., Kunisato, Y., Yoshino, A., Ueda, K., Suzuki, S., & Yamawaki, S. (2017). Cognitive behavioral therapy changes functional connectivity between medial prefrontal and anterior cingulate cortices, Journal of Affective Disorders, 15(208), 610-614. [Link](https://www.sciencedirect.com/science/article/abs/pii/S0165032716305195)

### 2016

- Takagaki, K., Okamoto, Y., Jinnin, R., Mori, A., Nishiyama, Y., Yamamura, T., Yokoyama, S., Shiota, S., Okamoto, Y., Miyake, Y., Ogata, A., Kunisato, Y., Shimoda, H., Kawakami, N., Furukawa, A. T., & Yamawaki. (2016). Behavioral activation for late adolescents with subthreshold depression: a randomized controlled trial. European Child & Adolescent Psychiatry, 25(11), 1171-1182. [Link](https://link.springer.com/article/10.1007/s00787-016-0842-5)

- Hasegawa, A., Nishimura, H., Matsuda, Y., Kunisato, Y., Morimoto, H., & Adachi, M. (2016). Is trait rumination associated with the ability to generate effective problem solving strategies? Utilizing two versions of the Means-Ends Problem-Solving Test. Journal of Rational-Emotive & Cognitive-Behavior Therapy, 34, 14-30. [Link](https://pubmed.ncbi.nlm.nih.gov/26900257/)

### 2015

- Tsukue, R., Okamoto, Y., Yoshino, A., Kunisato, Y., Takagaki, K., Takebayashi, Y., Tanaka, K., Konuma, K., Tsukue, I., & Yamawaki, S. (2015) Do Individuals with Alcohol Dependence Show Higher Unfairness Sensitivity? The Relationship Between Impulsivity and Unfairness Sensitivity in Alcohol-Dependent Adults. Alcoholism, clinical and experimental research, 39(10), 2016-2021.  [Link](https://onlinelibrary.wiley.com/doi/abs/10.1111/acer.12832)

- Nishiyama,Y., Okamoto, Y., Kunisato, Y., Okada, G., Yoshimura, S., Kanai, Y., Yamamura, T., Yoshino, A., Jinnin, R., Takagaki, K., Onoda, K., & Yamawaki, S. (2015) fMRI Study of Social Anxiety during Social Ostracism with and without Emotional Support. PLoS ONE 10(5): e0127426. [Link](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0127426)

### 2014

- Machino, A., Kunisato, Y., Matsumoto, T., Yoshimura, S., Ueda, K., Yamawaki, Y., Okada, G., Okamoto, Y., & Yamawaki, S (2014) Possible involvement of rumination in gray matter abnormalities in persistent symptoms of major depression: an exploratory magnetic resonance imaging voxel-based morphometry study, Journal of Affective Disorders, 168, 229-235. [Link](https://www.sciencedirect.com/science/article/abs/pii/S0165032714003917?via%3Dihub)

- Yoshino, A., Okamoto, Y., Kunisato, Y., Yoshimura, Y., Jinnin, R., Hayashi, Y., Kobayakawa, M., Doi, M., Oshita, K., Nakamura, R., Tanaka, K., Yamashita, H., Kawamoto, M., & Yamawaki, S. (2014). Distinctive spontaneous regional neural activity in patients with somatoform pain disorder: A resting-state fMRI preliminary study, Psychiatry Research: Neuroimaging, 221(3), 246-248. [Link](https://www.sciencedirect.com/science/article/abs/pii/S0925492713003417?via%3Dihub)

- Toki, S., Okamoto, Y., Onoda, K., Matsumoto, T., Yoshimura, S., Kunisato, Y., Okada, G., Shishida, K., Kobayakawa, M., Fukumoto, T., Machino, A., Inagaki, M., Yamawaki, S. (2014). Hippocampal activation during associative encoding of word pairs and its relation to symptomatic improvement in depression: A functional and volumetric MRI study. Journal of Affective Disorder, 152, 462-467. [Link](https://www.sciencedirect.com/science/article/abs/pii/S0165032713005892?via%3Dihub)

### 2013

- Yoshimura, S., Okamoto, Y., Onoda, K., Matsunaga, M., Okada, G., Kunisato, Y., Yoshino, A., Ueda, K., Suzuki, S., & Yamawaki, S. (2013). Cognitive behavioral therapy for depression changes medial prefrontal and ventral anterior cingulate cortex activity associated with self-referential processing. Social Cognitive and Affective Neuroscience, 9(4), 487-493.  [Link](https://academic.oup.com/scan/article/9/4/487/1629777)

- Takagaki, K., Okajima, I., Kunisato, Y., Nakajima, S., Kanai, Y., Ishikawa, S., & Sakano, Y. (2013). Preliminary assessment of the behavioral activation model in Japanese undergraduate students. Psychological Reports, 112(1), 47-59.  [Link](https://journals.sagepub.com/doi/abs/10.2466/02.15.21.PR0.112.1.47-59)

- Koseki, S., Noda., T., Yokoyama, S., Kunisato, Y., Ito, D., Suyama, H., Matsuda, T., Sugimura, Y., Ishihara, N., Shimizu, Y., Nakazawa, K., Yoshida, S., Arima, K., Suzuki,S. (2013). The relationship between positive and negative automatic thought and activity in the prefrontal and temporal cortices: A multi-channel near-infrared spectroscopy (NIRS) study, Journal of Affective Disorders, 151(1), 352-359. [Link](https://www.sciencedirect.com/science/article/pii/S0165032713004461)

### 2012

- Kinoshita, A., Okamoto, Y., Okada, G., Demoto, Y., Kunisato, Y., Yoshimura, S., Onoda,K., Kamachi, M., Yamawaki, S. (2012) Sex differences in neural activation to ambiguous facial expression in happy and sad context, Perceptual and Motor Skills, 115(2), 349-359. [Link](https://journals.sagepub.com/doi/abs/10.2466/22.24.PMS.115.5.349-359)

- Kunisato, Y., Okamoto, Y., Ueda, K., Onoda, K., Okada, G., Yoshimura, S., Suzuki, S., Samejima, K., Yamawaki, S. (2012) Effects of depression on reward based decision making and variability of action in probabilistic learning, Journal of Behavior Therapy and Experimental Psychiatry, 43, 1088-1094.  [Link](https://www.sciencedirect.com/science/article/abs/pii/S000579161200047X?via%3Dihub)

- Yoshino, A., Okamoto, Y., Onoda, K., Shishida, K., Yoshimura, S., Kunisato, Y., Demoto, Y., Okada, G., Toki, S., Yamashita, H., Yamawaki, S. (2012) Sadness enhances the experience of pain and affects pain-evoked cortical activities: An MEG study, The Journal of Pain,13, 628-635.  [Link](Sadness enhances the experience of pain and affects pain-evoked cortical activities: An MEG study)

- Demoto, Y., Okada, G., Okamoto, Y., Kunisato, Y., Aoyama, S., Onoda, K., Munakata, A., Nomura, M., Tanaka, S. C., Schweighofer, N., Doya, K., Yamawaki, S. (2012) Neural And Personality Correlates of Individual Differences Related to the Effects of Acute Tryptophan Depletion on Future Reward Evaluation, Neuropsychobiology, 65, 55-64.  [Link](https://www.karger.com/Article/Abstract/328990)

### 2011

- Okada, G., Okamoto, Y., Kunisato, Y., Aoyama, S., Nishiyama, Y., Yoshimura, S., Onoda, K., Toki, S., Yamashita, H., & Yamawaki, S. (2011) The effect of negative and positive emotionality on associative memory: An fMRI study, PLoS ONE, 6(9),e24862.  [Link](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0024862)

- Onoda, K., Okamoto, Y., Kunisato, Y., Aoyama, S., Shishida, K., Okada, G., Tanaka, SC., Schweighofer, N., Yamaguchi, S., Doya, K., & Yamawaki, S. (2011) Inter-individual discount factor differences in reward prediction are topographically associated with caudate activation, Experimental Brain Research, 212(4), 593-601.  [Link](https://link.springer.com/article/10.1007/s00221-011-2771-3)

- Kunisato, Y., Okamoto, Y., Okada, G., Aoyama, S., Nishiyama, Y., Onoda, K., Yamawaki, S. (2011)Personality traits and the amplitude of spontaneous low-frequency oscillations during resting state, Neuroscience Letters, 492(2), 109-113.  [Link](https://www.sciencedirect.com/science/article/abs/pii/S0304394011001133)

- Kunisato,Y., Okamoto, Y., Okada, G., Aoyama, S., Demoto, Y., Munakata, A., Nomura, M., Onoda, K., Yamawaki, S. (2011) Modulation of default - mode network activity by acute tryptophan depletion is associated with mood change：A resting state functional magnetic resonance imaging study. Neuroscience Research, 69, 129-134. [Link](https://www.sciencedirect.com/science/article/pii/S016801021002835X?via%3Dihub)

### 2010

- Matsunaga, M., Okamoto, Y., Suzuki, S., Kinoshita, A., Yoshimura, S., Yoshino, A., Kunisato, Y. & Yamawaki, S. (2010). Psychosocial functioning in patients with treatment-resistant depression after group cognitive behavioral therapy. Bmc Psychiatry, 10(1), 22. [Link](https://bmcpsychiatry.biomedcentral.com/articles/10.1186/1471-244X-10-22)

- Yoshino, A., Okamoto, Y., Onoda, K., Yoshimura, S., Kunisato, Y., Demoto, Y., Okada, G, & Shigeto, Y. (2010). Sadness enhances the experience of pain via neural activation in the anterior cingulate cortex and amygdala: An fMRI study. Neuroimage, 50(3), 1194-1201. [Link](https://www.sciencedirect.com/science/article/abs/pii/S1053811909012658)