### 2020-07-23 team.Rmd

---
title: "Team"
---

### Principal investigator

- [Yoshihiko Kunisato](yoshihiko-kunisato.html)

### Postdoctoral Fellow

- [Asako Toyama](https://t-asa.github.io/homepage/index.html)

### Ph.D. students

- [Keita Somatori](https://ksomatori.github.io/homepage/index.html)

### Master Course students

- Keisuke Inoue
- Satoshi Yamazaki

### Undergraduate students

- Yukiko Saito
- Jihee Woo
- Ryota Tanaka
- Daichi Igarashi
- Naoki Sawamura
- Shu Ensei
- Koji Komine