### 2021-07-24 team.Rmd

---
title: "Team"
---

### Principal investigator

- [Yoshihiko Kunisato](yoshihiko-kunisato.html)

### Postdoctoral Fellow

- [Asako Toyama](https://t-asa.github.io/homepage/index.html)

### Ph.D. students

- [Keita Somatori](https://ksomatori.github.io/homepage/index.html)

### Master Course students

- Keisuke Inoue
- Satoshi Yamazaki
- Yukiko Saito

### Undergraduate students
- Koji Komine
- Naoki Sawamura
- Ensei Shu
- Kaichu Kyo
- Yura Muraguchi
- Mayu Okano
- Masataka Sumida
- Kana Warashina